---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# fauxnaif

***faux-naïf*** (/ˌfoʊ.naɪˈif/ *FOH-niy-EEF*): a person who pretends to be simple or innocent

**`fauxnaif`**: an R package for pretending values are `NA`s

## Overview

fauxnaif provides a replacement for `dplyr::na_if()`.
Unlike [**dplyr**](https://github.com/tidyverse/dplyr)'s `na_if()`, it allows you
to specify multiple values to be replaced with `NA` using a single function.
Load **fauxnaif** after **dplyr** (or use a conflict manager like
[**conflicted**](https://github.com/r-lib/conflicted)) to use the extended
`na_if()` functionality.
Alternatively, use `na_if_in()` to avoid conflicts with **dplyr**.

## Installation

```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("rossellhayes/fauxnaif")
```

## Usage

```{r, message = FALSE}
library(dplyr)
library(fauxnaif)

na_if(1:5, 2, 4)

dplyr::starwars %>%
  select(name, hair_color) %>%
  mutate(mutated_hair_color = na_if(hair_color, "unknown", "none"))
```
