---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# fauxnaif
<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Build Status](https://travis-ci.org/rossellhayes/fauxnaif.svg?branch=master)](https://travis-ci.org/rossellhayes/fauxnaif)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/rossellhayes/fauxnaif?branch=master&svg=true)](https://ci.appveyor.com/project/rossellhayes/fauxnaif)
[![Codecov test coverage](https://codecov.io/gh/rossellhayes/fauxnaif/branch/master/graph/badge.svg)](https://codecov.io/gh/rossellhayes/fauxnaif?branch=master)
[![GitHub License](https://img.shields.io/github/license/rossellhayes/fauxnaif?color=blueviolet)](https://github.com/rossellhayes/fauxnaif/blob/master/LICENSE)
<!-- badges: end -->

***faux-naïf*** ([/ˌfoʊ.naɪˈif/](https://en.wikipedia.org/wiki/Help:IPA/English)): a person who pretends to be simple or innocent

**`fauxnaif`**: an R package for simplifying data by pretending values are `NA`

## Overview

fauxnaif provides a replacement for `dplyr::na_if()`.
Unlike [**dplyr**](https://github.com/tidyverse/dplyr)'s `na_if()`, it allows you
to specify multiple values to be replaced with `NA` using a single function.
Load **fauxnaif** after **dplyr** (or use a conflict manager like
[**conflicted**](https://github.com/r-lib/conflicted)) to use the extended
`na_if()` functionality.
Alternatively, use `faux_na_if()` to avoid conflicts with **dplyr**.

## Installation

```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("rossellhayes/fauxnaif")
```

## Usage

```{r, message = FALSE}
library(dplyr)
library(fauxnaif)

na_if(1:10, 2, 4:5, c(7, 9))

na_if(1:10, ~ . >= 8)

dplyr::starwars %>%
  select(name, hair_color) %>%
  mutate(mutated_hair_color = na_if(hair_color, "unknown", "none"))
```
